---
title: "Lab01"
author: "Janish Parikh"
date: "05/08/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---
Link for github depository:https://github.com/Ansh-Ji/CS-605-Data-Analytics-in-R/tree/master/Lab1


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library('ggplot2')
```
Question-(01)
The numbers of blocked intrusion attempts on each day during the first two weeks
of the month were 56, 47, 49, 37, 38, 60, 50, 43, 43, 59, 50, 56, 54, 58.
After the change of firewall settings, the numbers of blocked intrusions during 
the next 20 days were
53, 21, 32, 49, 45, 38, 44, 33, 32, 43, 53, 46, 36, 48, 39, 35, 37, 36, 39,45.
Comparing the number of blocked intrusions before and after the change,

```{r}
before<-c(56, 47, 49, 37, 38, 60, 50, 43, 43, 59, 50, 56, 54, 58)
after<-c(53, 21, 32, 49, 45, 38, 44, 33, 32, 43, 53, 46, 36, 48, 39, 35, 37, 36,
         39,45)
```

(a) construct side-by-side stem-and-leaf plots
```{r}
#stem(before)
#stem(after)
library('aplpack')
stem.leaf.backback(before,after)#use package aplpack(another plot package)

```

(b) compute the five-point summaries  
```{r}
summary(before)
summary(after)
```

construct parallel boxplots

```{r,echo= FALSE}
boxplot(before,after,main = "Parallel Boxplots",
        xlab=("no of blocked intrusion attempts"),
        names=c('Before','After'),col = c("orange","red"),
        border = "brown",horizontal = T)
```

(c) comment on your findings.

The mean no of  blocked intrusion attempts has dropped by 9.80.

The no of blocked intrusion attempts have significantly reduced after the new firewall settings. It has proved to be beneficial.

Question-(02) 

A network provider investigates the load of its network. The number of concurrent users is recorded at fifty locations (thousands of people) 17.2 22.1 18.5 17.2 18.6 14.8 21.7 15.8 16.3 22.8 24.1 13.3 16.2 17.5 19.0 23.9 14.8 22.2 21.7 20.7 13.5 15.8 13.1 16.1 21.9 23.9 19.3 12.0 19.9 19.4 15.4 16.7 19.5 16.2 16.9 17.1 20.2 13.4 19.8 17.7 19.7 18.7 17.6 15.9 15.2 17.1 15.0 18.8 21.6 11.9 

```{r}
user<-c(17.2, 22.1, 18.5, 17.2, 18.6, 14.8, 21.7, 15.8, 16.3, 22.8,  24.1, 13.3, 
        16.2, 17.5, 19.0,23.9, 14.8, 22.2, 21.7, 20.7 ,13.5 ,15.8, 13.1, 16.1, 
        21.9, 23.9, 19.3, 12.0, 19.9, 19.4, 15.4, 16.7, 19.5, 16.2, 16.9, 17.1, 
        20.2, 13.4, 19.8, 17.7, 19.7 ,18.7, 17.6, 15.9, 15.2 ,17.1 ,15.0, 18.8, 
        21.6 ,11.9)
```

(a) Compute the sample mean, variance, and standard deviation of the number of concurrent users. 

```{r}
explore<-function(x){
  data<-c("Mean"=mean(x),"Median"=median(x), 
          "Standard Deviation" = sd(x))
  return(data)
}
explore(user)
```

(b) Estimate the standard error of the sample mean. 
```{r}
#function to calculate deviation from any point
#default value set to mean
standard_deviation<-function(x=numeric(),y=mean(x)){
  deviation<-sqrt(sum((x-y)^2/(length(x)-1))) 
  return(deviation)}

#function to calculate standard error but of any point
standard_error<-function(x=numeric(),y=mean(x)){
  error<-(standard_deviation(x,y)/sqrt(length(x)))
  return(error)
}
standard_error(user)
library('psych')
describe(user,IQR=T)

```

(c)Compute the five-point summary and construct a boxplot.

```{r}
summary(user)
```

```{r , echo=FALSE}
boxplot(user,  
        main="Network Load",
        xlab = "Location",
        ylab = "No of concurrent users in thousands",
        col = "orange"
        )
```

(d) Compute the interquartile range. Are there any outliers?  
```{r}
# IQR is measure of statistical dispersion IQR=Q3-Q1
Q1<-quantile(user,0.25)
Q3<-quantile(user,0.75)
IQR<-Q3-Q1
print(IQR)
#Used 1.5 IQR
left<- (Q1-(1.5*IQR))
right<- (Q3+(1.5*IQR))
outliers<-c(user[user <left],user[user>right])
print(outliers)
```

There are no Outliers

(e)It is reported that the number of concurrent users follows approximately Normal distribution. Does the histogram support this claim? 
```{r , echo=FALSE}
#function to plot histogram
plot_histogram<-function(x){
  p<-ggplot(data=data.frame(data=x),aes(x=data))+
    geom_histogram(aes(y=stat(ndensity)),bins=10)+
    geom_density(aes(y=stat(scaled)),fill = "red",alpha = 0.1)
  plot(p)
}
plot_histogram(user)
```

The distribution looks like a normal distribution

Question-(03) 

Consider three data sets.
(1) 19, 24, 12, 19, 18, 24, 8, 5, 9, 20, 13, 11, 1, 12, 11, 10, 22, 21, 7, 16,
15, 15, 26, 16, 1, 13, 21, 21, 20, 19
(2) 17, 24, 21, 22, 26, 22, 19, 21, 23, 11, 19, 14, 23, 25, 26, 15, 17, 26,
21, 18, 19, 21, 24, 18, 16, 20, 21, 20, 23, 33
(3) 56, 52, 13, 34, 33, 18, 44, 41, 48, 75, 24, 19, 35, 27, 46, 62, 71, 24,
66, 94, 40, 18, 15, 39, 53, 23, 41, 78, 15, 35

```{r}
dataset1<-c(19, 24, 12, 19, 18, 24, 8, 5, 9, 20, 13, 11, 1, 12, 11, 10, 22, 21, 7, 16,
        15, 15, 26, 16, 1, 13, 21, 21, 20, 19)
dataset2<-c(17, 24, 21, 22, 26, 22, 19, 21, 23, 11, 19, 14, 23, 25, 26, 15, 17, 26,
        21, 18, 19, 21, 24, 18, 16, 20, 21, 20, 23, 33)
dataset3<-c(56, 52, 13, 34, 33, 18, 44, 41, 48, 75, 24, 19, 35, 27, 46, 62, 71, 24,
        66, 94, 40, 18, 15, 39, 53, 23, 41, 78, 15, 35)
```

(a) For each data set, draw a histogram and determine whether the
distribution is rightskewed, left-skewed, or symmetric
```{r , echo=FALSE}
#function to plot histogram
plot_histogram<-function(x){
  p<-ggplot(data=data.frame(data=x),aes(x=data))+
    geom_histogram(aes(y=stat(ndensity)),bins=10)+
    geom_density(aes(y=stat(scaled)),fill = "red",alpha = 0.1)
  plot(p)
}
plot_histogram(dataset1)
```
Left Skewed
```{r , echo=FALSE}
plot_histogram(dataset2)
```
Uniform
```{r, echo=FALSE}
#function to plot histogram
plot_histogram(dataset3)
```
Right Skewed

(b) Compute sample means and sample medians. 
```{r}
#function to calculate sample mean and median
sample_mean_median<-function(x){
  data<-c("Mean"=mean(x),"Median"=median(x))
  return(data)}
sample_mean_median(dataset1)
sample_mean_median(dataset2)
sample_mean_median(dataset3)
```
(b)Do they support your findings about skewness and symmetry? How?
```{r}
mean(dataset1)-median(dataset1)

```
As mean is less than median we can say it is left skewed
```{r}
mean(dataset3)-median(dataset3)

```
As mean is greater than median it is right skewed

We can determine the same(not always correct) via the help of histogram
If the histogram has a longer left tail then we call it to be left skewed, same follows for right


Question-(04) 
The following data set represents the number of new computer accounts 
registered during ten consecutive days.
43, 37, 50, 51, 58, 105, 52, 45, 45, 10


```{r}
dataset<-c(43, 37, 50, 51, 58, 105, 52, 45, 45, 10)
```
(a) Compute the mean, median, quartiles, and standard deviation
```{r}
explore<-function(x){
  data<-c("Mean"=mean(x),"Median"=median(x), 
          "Standard Deviation" = sd(x),"Quartile-" =quantile(x))
  
  return(data)
}
explore(dataset)
```
(b) Check for outliers using the 1.5(IQR) rule.
```{r dataset}
###Below Q1-1.5*IQR  or above Q3+1.5*IQR, where Q1 and Q3 are the first and third 
#quartiles, respectively, of the variable distribution and IQR=Q3-Q1 is the 
#interquartile range.

IQR_outliers <- function(x) {
  Q1<-quantile(x,0.25)
  Q3<-quantile(x,0.75)
  IQR<-(Q3-Q1) #inter-quartile range
  left<- (Q1-(1.5*IQR))
  right<- (Q3+(1.5*IQR))
  outliers<-c(x[x <left],x[x>right])
  return(outliers)
}
outliers<-IQR_outliers(dataset)
print(outliers)
```

```{r , echo=FALSE}
boxplot(dataset,  
        main="Checking Outliers",
        col = "orange",
        horizontal = T
        )
```
The same outliers can be pointed out via a boxplot


(c) Delete the detected outliers 

```{r }
reconstructed_dataset<-dataset[!dataset%in%outliers]
print(reconstructed_dataset)
```
compute the mean, median, quartiles, and standard deviation again
```{r}
explore(reconstructed_dataset)
```

(d) Make a conclusion about the effect of outliers on basic descriptive statistics
We can do this by analyzing the changes in Mean, Median, Standard Deviation
```{r}
difference<-explore(dataset)-explore(reconstructed_dataset)
print(difference )
```
Quartile .50% or Median remains unchanged 
Standard Deviation decreases by 17.01943 

Mean is affected  
```{r }
#function to plot histogram
plot_histogram<-function(x){
  p<-ggplot(data=data.frame(data=x),aes(x=data))+
    geom_histogram(aes(y=stat(ndensity)),bins=10)+
    geom_density(aes(y=stat(scaled)),fill = "red",alpha = 0.1)
  plot(p)
}
plot_histogram(dataset)
```

```{r }
plot_histogram(reconstructed_dataset)
```

After removing the outliers we can see that the spread of data has decreased and the data is more closely distributed.

