---
title: "Lab02"
author: "Janish Parikh"
date: "12/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('dplyr')
library('ggplot2')
library('ggpubr')
library("rmutil")
```
#Question-01 Central limit theorem:
(A) Generate mean subtracted X1, X2, ..., XN
uniformly distributed, i.e., U[0, 1], IID random variables. Generate at
least 10000 samples for each of the random variables and let N = 25.
```{r}
set.seed(1)
N<-25
randomvariables<-matrix(data=rep(0,25000), nrow = 10000, ncol = N)
for (i in c(1:N)){
  randomvariables[,i] <-runif(10000, 0, 1) 
  randomvariables[,i]<-randomvariables[,i]-mean(randomvariables[,i])
} 
colnames(randomvariables)<-c("X1","X2","X3","X4",'X5',"X6","X7",'X8',"X9","X10",
                             "X11","X12","X13",'X14',"X15","X16","X17",
                             "X18",'X19',"X20","X21","X22","X23","X24","X25")
samplespace<-tbl_df(data=randomvariables)
head(samplespace)
```

starting from X1 start adding the random variables, i.e.,

Compute Yk= sum(Xi....Xk) Plot the PDFs for each of the Yk.

```{r}
plot_histogram<-function(x){
  p<-ggplot(data=data.frame(data=x),aes(x=data))+
    geom_histogram(aes(y=stat(ndensity)),bins=10)+
    geom_density(aes(y=stat(scaled)),fill = "red",alpha = 0.1)
  plot(p)
}
```

```{r}
pdf<-function(samplespace,ytable, N =25){
ytable<-samplespace
colnames(ytable)<-gsub("X","Y",colnames(samplespace))
for (i in c(1:N)){
  s1<-select(samplespace,c(1:i)) 
  ytable[,i]<-rowSums(s1)
} 
head(ytable)
x<-list(ytable)
 list<-sapply(x[[1]], plot_histogram)}
pdf(samplespace)
```

(B) Repeat the above experiment for mean subtracted independent random variables, with different distributions, e.g., Uniform, Laplacian, etc.

Compare your results with the above experiment.

```{r}
set.seed(1)
for (i in c(1:N)){
  if(i%%2==0){
  randomvariables[,i] <-rlaplace(10000, 0.5, 0.1) 
  } else{
    randomvariables[,i]<-runif(10000,0,1)
  }
  randomvariables[,i]<-randomvariables[,i]-mean(randomvariables[,i])
} 
samplespacemixed<-tbl_df(data=randomvariables)

pdf(samplespacemixed,ytable = ytablemixed)
```

(C) Finally, perform the same experiment for dependent random vari-
ables and compare your results. For example, X1 ∼ U[−1, 1]; X2 =2X1; X3 = X1 + X2 + 3, and likewise.
```{r}
set.seed(1)
randomvariables[,1]<-runif(10000,-1,1)
randomvariables[,2]<-2*randomvariables[,1]
randomvariables[,3]<-randomvariables[,1]+randomvariables[,2]
for (i in c(4:N)){
  if(i%%2==0){
  randomvariables[,i] <-randomvariables[,2] +i 
  } else{
    randomvariables[,i]<-randomvariables[,3] + i 
  }
  randomvariables[,i]<-randomvariables[,i]-mean(randomvariables[,i])
} 
dependentrv<-tbl_df(data=randomvariables)

pdf(dependentrv,ydependent)
```

```{r}
rm(i,N,samplespace,samplespacemixed,dependentrv,randomvariables)
```

#Q. 2: Sampling distribution of mean: 
A manufacturing process produces cylindrical component parts for the
automobile industry. It is required to produce the parts with a mean
diameter of 5.0 mm. It is known that the population standard deviation
is σ = 0.1 mm. The engineer involved conjectures that the population
mean is 5.0 mm.

```{r}
mean = 5.0 #a+b/2
sd = 0.1 #variance = (b-a)^2/12
b=round(5+0.1*sqrt(3),2)
a=10-b
```

A)An experiment is conducted in which 100 parts produced are selected
randomly and the diameter measured in each. Generate uniformly distributed 100 random numbers to represent the diameter using available
information and compute a sample average diameter. Does this sample
information appear to support or refute the engineer’s conjecture? Draw
necessary plots.
```{r}
set.seed(1)
diameter<-runif(100,min=a,max=b)
sampleavgdiameter<-mean(diameter)
samplevariance<-var(diameter)
plot_histogram(diameter)
zs<-(sampleavgdiameter-mean)/sd #zscore
pnorm(zs) #Getting the probability value

```
(B)Generate the sampling distribution of mean by repeating the above
random experiment several times. Based on the sample mean obtained
thus, comment on the engineer’s conjecture? Draw necessary plots.
```{r}
set.seed(1)
samplingmean<-function(s,n,a=4.83,b=5.17){ #s is no of simulation and n is the no of trials
  meantable<-matrix(data=NA, nrow = s,ncol=3)
  colnames(meantable)<-c('Mean','Zscore','PVAL')
  for (i in c(1:s)){
    diameter<-diameter<-runif(n,min=a,max=b)
    meantable[i,1]<-mean(diameter)
    meantable[i,2]<-(mean(diameter)-(a+b)/2)/(0.1/sqrt(n))
    meantable[i,3]<-pnorm(meantable[i,2])
  }
  plot_histogram(meantable[,1])
  return(meantable)
}
data<-samplingmean(100,100)
summary(data)
samplemean<-mean(data[,1])
zs<-(samplemean-mean)/(sd/sqrt(100)) #Z score
print(pnorm(zs))
```


(C)Repeat the part (B) by repeating the random experiment for 10000
parts. Based on the sample mean obtained thus, comment on the engineer’s conjecture? 
Draw necessary plots.
```{r}
set.seed(1)
data<-samplingmean(10000,100)
samplemean<-mean(data[,1])
summary(data)
  zs<-(samplemean-5.0)/(0.1/sqrt(10000))
  print(pnorm(zs))
```

```{r}
rm(data,a,b,diameter,mean,sampleavgdiameter,samplemean,samplevariance,sd,zs)
```
#Q-3Sampling distribution of the difference between two means:
Consider two populations with certain probability distributions P1(Ω)
and P2(Ω). Draw two independent random samples of size n1 and n2
from each population, respectively. Now, generate the sampling distributions 
of differences of the two means under following cases, and
comment on the Normal approximation of the sampling distribution of
difference between two means

(A) For P1(Ω) ∼ N(5, 1) and P2(Ω) ∼ N(3, 1), and for n1 < 30 and n2 < 30.
```{r}
set.seed(1)
library('latex2exp')
n =20
m1 = 5
s1 = 1
m2 = 3
s2 = 1
num_exp = 500
mean_diff = 0 * c(1:n)
for(i in c(1:n)){
  temp1 = rnorm(num_exp, mean = m1, sd = s1)
  temp2 = rnorm(num_exp, mean = m2, sd = s2)
  mean_diff[i] = mean(temp1) - mean(temp2)
}
p1 <- ggplot(data = data.frame(x = mean_diff), aes(x = x))+
  geom_histogram(aes(y = stat(ndensity)), bins = 15)+
  geom_density(aes(y = stat(scaled)), fill = "blue", alpha = 0.15)+
  labs(x = TeX("$\\bar{X}_1 - \\bar{X}_2$"), y = "Normalized Frequency",
       title = "Case 1")+
  theme(plot.title = element_text(hjust = 0.5), plot.margin = unit(c(0.25, 1, 0.25, 0.15), "cm"))
plot(p1)

```


(B)Repeat part (A) for non-normal populations while keeping other settings the same.
```{r}
set.seed(1)
n = 20
m1 = 5
s1 = 1
m2 = 3
s2 = 1
num_exp = 500
mean_diff = 0 * c(1:n)
for(i in c(1:n)){
  temp1 = runif(num_exp, min = 2, max = 8)
  temp2 = runif(num_exp, min = 1, max = 5)
  mean_diff[i] = mean(temp1) - mean(temp2)
}
p2 <- ggplot(data = data.frame(x = mean_diff), aes(x = x))+
  geom_histogram(aes(y = stat(ndensity)), bins = 12)+
  geom_density(aes(y = stat(scaled)), fill = "red", alpha = 0.15)+
  labs(x = TeX("$\\bar{X}_1 - \\bar{X}_2$"), y = "Normalized Frequency",
       title = "Case 2")+
  theme(plot.title = element_text(hjust = 0.5))
plot(p2)
```

(C) For P1(Ω) ∼ N(5, 1) and P2(Ω) ∼ N(3, 1), and for n1 ≥ 30 and n2 ≥ 30.
```{r}
set.seed(1)
n = 500
m1 = 5
s1 = 1
m2 = 3
s2 = 1
num_exp = 500
mean_diff = 0 * c(1:n)
for(i in c(1:n)){
  temp1 = rnorm(num_exp, mean = m1, sd = s1)
  temp2 = rnorm(num_exp, mean = m2, sd = s2)
  mean_diff[i] = mean(temp1) - mean(temp2)
}
p3 <- ggplot(data = data.frame(x = mean_diff), aes(x = x))+
  geom_histogram(aes(y = stat(ndensity)), bins = 25)+
  geom_density(aes(y = stat(scaled)), fill = "yellow", alpha = 0.15)+
  labs(x = TeX("$\\bar{X}_1 - \\bar{X}_2$"), y = "Normalized Frequency",
       title = "Case 3")+
  theme(plot.title = element_text(hjust = 0.5), plot.margin = unit(c(0.25, 1, 0.25, 0.15), "cm"))
plot(p3)

```

(D) Repeat part (C) for non-normal populations while keeping other settings the same.
```{r}
set.seed(1)
n = 500
m1 = 5
s1 = 1
m2 = 3
s2 = 1
num_exp = 500
mean_diff = 0 * c(1:n)
for(i in c(1:n)){
  temp1 = runif(num_exp, min = 2, max = 8)
  temp2 = runif(num_exp, min = 1, max = 5)
  mean_diff[i] = mean(temp1) - mean(temp2)
}
p4 <- ggplot(data = data.frame(x = mean_diff), aes(x = x))+
  geom_histogram(aes(y = stat(ndensity)), bins = 30)+
  geom_density(aes(y = stat(scaled)), fill = "orange", alpha = 0.15)+
  labs(x = TeX("$\\bar{X}_1 - \\bar{X}_2$"), y = "Normalized Frequency",
       title = "Case 4")+
  theme(plot.title = element_text(hjust = 0.5))
plot(p4)
```

Draw all necessary plots for detail illustration.

```{r}
ggarrange(p1, p2, p3, p4, nrow = 2, ncol = 2)
```

```{r}
rm(i,m1,m2,mean_diff,n,s1,s2,temp1,temp2,p1,p2,p3,p4,num_exp)
```